---
title: "Defusing R Expressions with rlang"
author: "Gary Hutson"
date: "14/10/2021"
output:
  html_document:
    theme: lumen
    highlight: tango
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rlang)
```

## Types of defused expressions

- Calls, like f(1, 2, 3), 1 + 1, or ..., represent the action of calling a function to compute a new value, such as a vector.

- Symbols, like x or df, represent named objects. When the object pointed to by the symbol was defined in a function or in the global environment, we call it an environment-variable. When the object is a column in a data frame, we call it a data-variable.


### Create new symbol

```{r new_expression}
df <- NHSRdatasets::stranded_data

# Call the symbol - get the object passed and not the result
expr(df$stranded.label)

# Create a call representing the computation of the mean
expr(mean(df$stranded.label, na.rm=TRUE))


```

Defusing is not the only way to create defused expressions. You can also assemble them from data:

```{r defuse_from_data}
# Assemble a symbol from a string
var <- "stranded_data"
sym(var)
# Assemble a call from strings, symbols, and other objects
call("mean", sym(var), na.rm = TRUE)

```

# Defusing function arguments

There are two points of view when it comes to defusing an expression:

1. You can defuse expressions that you supply with **`expr()`**. This is one way of creating symbols and calls (see previous section).

2. You can defuse the expressions supplied by the user of your function with the operators starting with en like **`ensym()`**, **`enquo()`** and their plural variants. They defuse function arguments.

## Comparison to base R

- The defusing operator **`expr()`** is similar to `__quote()__`. Like **`bquote()`**, it allows forcing evaluation of parts of an expression.
- The plural variant exprs() is similar to alist().
- The argument-defusing operator enquo() is similar to substitute().


```{r capture_expression}

expr(i_passed_this)

# I can pass more I just need exprs

exprs(i, passed, this, as, multiple, symbols)
#Returns a list


```

To capture user expressions, most useful when designing functions, we just prefix with en:

```{r capturing_expression}

capture_multiple_expression_from_user <- function(some_arg, ...){
  user_exprs <- enexprs(some_arg, ...)
  return(user_exprs)
}

# User function
capture_multiple_expression_from_user(this, is, gary, these, were, variables, they, would, be, passed,
                                      through, the, myfunction, tidystyle)



```

Ensym and ensyms provide additional type checking to ensure the user calling your function has supplied bare object
names:

```{r ensym}
sym_inputs <- function(...){
  user_symbols <- ensyms(...)
  user_symbols
}

#sym_inputs(say(hello))
#This will generate an error

sym_inputs(hello, "bonjour")


```

## Unquoting and quasi-modo

All these quoting functions have quasiquotation support. This means that you can unquote (evaluate and inline) part of the captured expression:


```{r unquoting}
what <- sym("bonjour")
what

# Get the variable name passed
expr(say(what))

# Get the return value
expr(say(!!what))

# return a composite function
expr(say({what}))


```

Unquoting multiple:

```{r unquoting_multiple}

capture_multiple_expression_from_user(say(!!what), !!what, what)
# First is the unquoted variable
# Second is the value passed to what
# Third is the unquoted expression, meaning the actual value passed


```

# Quosures

These are slightly different to what we have learned, as quosures capture the expression and the environment.

```{r quosure}

my_quo <- quo(letters)
my_quo

# Get the expression
get_expr(my_quo)
# Get the environment
get_env(my_quo)

# Evaluate the captured expression
eval_tidy(my_quo)

# They have the nice property that you can pass them around from
# context to context (that is, from function to function) and they
# still evaluate in their original environment:

multiply_expr_by_10 <- function(expr){
  expr <- enquo(expr) #Capture the user expression as symbol
  
  local_ten <- 10 # Set up variable local to function otherwise global <<-
  
  quo(!!expr * local_ten) #Quote the unquoted user supplied expression to get the value * ten
}

```
The next step would be to evaluate this:

```{r evalue_captued_expression}
quo <- multiply_expr_by_10(2+3)
# This will still be a quosure and it will contain the envrionment it is evaluated in
quo
# Finally, we would just then evaluate the quosure
eval_tidy(quo)


```




